---
title: 'Final Project, P5: Replications'
author: "Nathaniel Evans, ALfonso Poire, Wei-Chun Lin"
date: "Dec 5, 2018"
output:
  html_document: 
    theme: flatly 
    highlight: kate
    toc: true
    toc_depth: 2
    toc_float : true
    smooth_scroll: true
    number_sections : false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(infer)
library(dplyr)
library(tidyr)
library(ggplot2)
library(GGally)
library(tidyverse)
library(DescTools)
```

# OVERVIEW

### PAPER: [Hair cortisol levels, psychological stress and psychopathological symptoms as predictors of postpartum depression](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0182817) 
### ORIGINAL DATA: [sav file](https://figshare.com/articles/PPD_hairCortisol_PlosOne_sav/5255848/1) 
### TIDY DATA: [csv file](https://github.com/OHSU-Math630/FINAL_PROJECT/tree/master/data/tidy_data.csv)
### CODE BOOK: [knitted html](https://github.com/OHSU-Math630/FINAL_PROJECT/tree/master/postpartum_depression_prediction/P2/group7_codebook.html)

#### The purpose of this R markdown script is to explore the data provided by this paper. If desired, all materials can be seen at our [github repository.](https://github.com/OHSU-Math630/FINAL_PROJECT) 

## Paper Abstract   

Postpartum depression affects a huge number of women and has detrimental consequences. Knowing the factors associated with postpartum depression during pregnancy can help its prevention. Although there is evidence surrounding behavioral or psychological predictors of postpartum depression, there is a lack of evidence of biological forecasters. The aim of this study was to analyze the sociodemographic, obstetric, and psychological variables along with hair cortisol levels during the first, second, and third trimesters of pregnancy that could predict postpartum depression symptoms. A sample of 44 pregnant women was assessed during 3 trimesters of pregnancy and the postpartum period using psychological questionnaires and hair cortisol levels. Participants were divided into 2 groups: a group with postpartum depression symptoms and a group with no postpartum depression symptoms. Results showed significant positive differences between groups in the first trimester regarding the Somatization subscale of the SCL-90-R (p < .05). In the second trimester, significant differences were found in the Somatization, Depression, Anxiety, and GSI subscales (p < .05). In the third trimester significant differences between both groups were found regarding pregnancy-specific stress. We found significant positive differences between groups regarding hair cortisol levels in the first and the third trimester. Hair cortisol levels could predict 21.7% of the variance of postpartum depression symptoms. In conclusion, our study provided evidence that psychopathological symptoms, pregnancy-specific stress, and hair cortisol levels can predict postpartum depression symptoms at different time-points during pregnancy. These findings can be applied in future studies and improve maternal care in clinical settings. [1]

# DATA ISSUES 
[Part 1 Rmd](https://github.com/OHSU-Math630/FINAL_PROJECT/blob/master/postpartum_depression_prediction/P1/DATA_QUAL_REVIEW.Rmd)
[Part 1 html](https://github.com/OHSU-Math630/FINAL_PROJECT/blob/master/postpartum_depression_prediction/P1/DATA_QUAL_REVIEW.html)

## Summary  
The main data quality issue we encountered was that the variable names were in Spanish, but additionally, we found that the metadata for categorical mapping, eg the original researchers codebook were only accessible through SPSS (.sav file). Becuase of this, it was problematic to interpret their data. Additionally, we considered the data highly curated, poor representation of the population, at risk of error from human entered data and potentially influenced by temporal components that were not accounted for. 

# CODE BOOK
[Part 2 Rmd](https://github.com/OHSU-Math630/FINAL_PROJECT/tree/master/postpartum_depression_prediction/P2/group_7_DataDelivery.Rmd)
[Part 2 html](https://github.com/OHSU-Math630/FINAL_PROJECT/tree/master/postpartum_depression_prediction/P2/group_7_DataDelivery.html)
[code book](https://github.com/OHSU-Math630/FINAL_PROJECT/tree/master/postpartum_depression_prediction/P2/group_7_codebook.html)

## Summary
The codebook went together fairly well after we got access to the .sav metadata through SPSS (thanks to wei-chun). The components to note on this part was that the datamaid produced a far cleaner codebook and was easier to use then the codebook library. 

# EDA
[Part 3 Rmd](https://github.com/OHSU-Math630/FINAL_PROJECT/blob/master/postpartum_depression_prediction/P3/group7_EDA.rmd)
[Part 3 html](https://github.com/OHSU-Math630/FINAL_PROJECT/blob/master/postpartum_depression_prediction/P3/group7_EDA.html)

## Summary
<lets add summary here>

# REPLICATION

## Summary 
<lets add summary here> 

## Load Tidy Dataset 

```{r}
load('../../data/tidy_data.Rdata')

data_eng <- read.csv('../../data/data_eng.csv')

pp_sad <- df_tidy # rename meaningful 

data_eng <- na.omit(data_eng)

data_eng$depreposparto <-as.factor(data_eng$depreposparto)
``` 
###Table & figures

##Table 1
```{r}
pp_sad %>%
  group_by(postpartum_depression) %>%
  summarise(sd_age = sd(age), mean_age = mean(age))

t.test(pp_sad$age ~ pp_sad$postpartum_depression, var.equal = T) 

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(nationality) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(nationality) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(employed) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(employed) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(	education_level) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(	education_level) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(sport) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(sport) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(pet) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(pet) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(dyed_hair) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(dyed_hair) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(first_pregnancy) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(first_pregnancy) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(wanted_pregnancy) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(wanted_pregnancy) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(pregnancy_method) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(pregnancy_method) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(previous_miscarriage) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(previous_miscarriage) %>%
   mutate(freq = n / sum(n)*100)

pp_sad %>%
  filter(postpartum_depression == 1) %>%
  count(fetus_sex) %>%
   mutate(freq = n / sum(n)*100)
pp_sad %>%
  filter(postpartum_depression == 2) %>%
  count(fetus_sex) %>%
   mutate(freq = n / sum(n)*100)

chisq.test(pp_sad$postpartum_depression, pp_sad$nationality, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$employed, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$education_level, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$sport, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$pet, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$dyed_hair, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$first_pregnancy, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$wanted_pregnancy, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$pregnancy_method, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$previous_miscarriage, correct=FALSE)
chisq.test(pp_sad$postpartum_depression, pp_sad$fetus_sex, correct=FALSE)

pp_sad %>%
  filter(epds >=10) %>%
  count
```

##Table2
```{r}
data_eng %>%
  group_by(depreposparto) %>%
  summarise(mean_PDQ1 = mean(PDQ1), sd_PDQ1=sd(PDQ1),  mean_SOMS1 = mean(SOMATIZATIONS1), sd_SOMS1 = sd(SOMATIZATIONS1), mean_DEP1 = mean(DEPRESSION1), sd_DEP1 = sd(DEPRESSION1), mean_anx1 = mean(ANXIETY1), sd_anx1 = sd(ANXIETY1), mean_GSI1 = mean(IGS1), sd_GSI1 = sd(IGS1))

data_eng %>%
  group_by(depreposparto) %>%
  summarise(mean_PDQ2 = mean(PDQ2), sd_PDQ2=sd(PDQ2),  mean_SOMS2 = mean(SOMATIZATIONS2), sd_SOMS2 = sd(SOMATIZATIONS2), mean_DEP2 = mean(DEPRESSION2), sd_DEP2 = sd(DEPRESSION2), mean_anx2 = mean(ANXIETY2), sd_anx2 = sd(ANXIETY2), mean_GSI2 = mean(IGS2), sd_GSI2 = sd(IGS2))

data_eng %>%
  group_by(depreposparto) %>%
  summarise(mean_PDQ3 = mean(PDQ3), sd_PDQ3=sd(PDQ3),  mean_SOMS3 = mean(SOMATIZATIONS3), sd_SOMS3 = sd(SOMATIZATIONS3), mean_DEP3 = mean(DEPRESSION3), sd_DEP3 = sd(DEPRESSION3), mean_anx3 = mean(ANXIETY3), sd_anx3 = sd(ANXIETY3), mean_GSI3 = mean(IGS3), sd_GSI3 = sd(IGS3))


t.test(data_eng$PDQ1 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$SOMATIZATIONS1 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$DEPRESSION1 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$ANXIETY1 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$IGS1 ~ data_eng$depreposparto, var.equal = T) 

t.test(data_eng$PDQ2 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$SOMATIZATIONS2 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$DEPRESSION2 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$ANXIETY2 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$IGS2 ~ data_eng$depreposparto, var.equal = T) 

t.test(data_eng$PDQ3 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$SOMATIZATIONS3 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$DEPRESSION3 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$ANXIETY3 ~ data_eng$depreposparto, var.equal = T) 
t.test(data_eng$IGS3 ~ data_eng$depreposparto, var.equal = T) 
##lm_PDQ1 <- lm(PDQ1 ~ depreposparto, data = data_eng)
##anova(lm_PDQ1)
```

##Figure 2
```{r}
tidy_SOM <- data_eng %>%
  select("X", "SOMATIZATIONS1", "SOMATIZATIONS2", "SOMATIZATIONS3", "depreposparto") %>%
  gather(trimester, somatization, SOMATIZATIONS1:SOMATIZATIONS3, -depreposparto)


ggplot(tidy_SOM, aes(x = trimester, y = somatization, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Somatization score", colour = "Postpartum Depression")+
  ggtitle('Somatization') +
  theme_bw() 
```

```{r}
tidy_obs <- data_eng %>%
  select("X", "OBSESSIONS.AND.COMPULSIONS1", "OBSESSIONS.AND.COMPULSIONS2", "OBSESSIONS.AND.COMPULSIONS3", "depreposparto") %>%
  gather(trimester, obsession, OBSESSIONS.AND.COMPULSIONS1:OBSESSIONS.AND.COMPULSIONS3, -depreposparto)


ggplot(tidy_obs, aes(x = trimester, y = obsession, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Obsession compulsion score", colour = "Postpartum Depression")+
  ggtitle('Obsession compulsion') +
  theme_bw() 
```

```{r}
tidy_is <- data_eng %>%
  select("X", "SENSITIVIDADINTERPERSONAL1", "INSTRUMENT.SENSITIVITY2", "INSTRUMENT.SENSITIVITY3", "depreposparto") %>%
  gather(trimester, INSTRUMENT, SENSITIVIDADINTERPERSONAL1:INSTRUMENT.SENSITIVITY3, -depreposparto)


ggplot(tidy_is, aes(x = trimester, y = INSTRUMENT, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Interpersonal senesitvity", colour = "Postpartum Depression")+
  ggtitle('Interpersonal senesitvity') +
  theme_bw() 
```

```{r}
tidy_ad <- data_eng %>%
  select("X", "DEPRESSION1", "DEPRESSION2", "DEPRESSION3", "depreposparto") %>%
  gather(trimester, depression, DEPRESSION1:DEPRESSION3, -depreposparto)


ggplot(tidy_ad, aes(x = trimester, y = depression, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Antenatal depression", colour = "Postpartum Depression")+
  ggtitle('Antenatal depression') +
  theme_bw() 
```

```{r}
tidy_anx <- data_eng %>%
  select("X", "ANXIETY1", "ANXIETY2", "ANXIETY3", "depreposparto") %>%
  gather(trimester, anxitety, ANXIETY1:ANXIETY3, -depreposparto)


ggplot(tidy_anx, aes(x = trimester, y = anxitety, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Anxitety", colour = "Postpartum Depression")+
  ggtitle('Anxitety') +
  theme_bw() 
```

```{r}
tidy_hos <- data_eng %>%
  select("X", "HOSTILIDAD1", "HOSTILITY2", "HOSTILIDAD3", "depreposparto") %>%
  gather(trimester, hostility, HOSTILIDAD1:HOSTILIDAD3, -depreposparto)


ggplot(tidy_hos, aes(x = trimester, y = hostility, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Hostility", colour = "Postpartum Depression")+
  ggtitle('Hostility') +
  theme_bw() 
```

```{r}
tidy_pa <- data_eng %>%
  select("X", "ANSIEDADFOBICA1", "ANSIEDADFOBICA2", "ANSIEDADFOBICA3", "depreposparto") %>%
  gather(trimester, p_anxiety, ANSIEDADFOBICA1:ANSIEDADFOBICA3, -depreposparto)


ggplot(tidy_pa, aes(x = trimester, y = p_anxiety, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Phobic anxiety", colour = "Postpartum Depression")+
  ggtitle('Phobic anxiety') +
  theme_bw() 
```

```{r}
tidy_pi <- data_eng %>%
  select("X", "IDEACIÃ.NPARANOIDE1", "IDEACIÃ.NPARANOIDE2", "IDEACIÃ.NPARANOIDE3", "depreposparto") %>%
  gather(trimester, paranoid, IDEACIÃ.NPARANOIDE1:IDEACIÃ.NPARANOIDE3, -depreposparto)


ggplot(tidy_pi, aes(x = trimester, y = paranoid, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Paranoid ideation", colour = "Postpartum Depression")+
  ggtitle('Paranoid ideation') +
  theme_bw() 
```

```{r}
tidy_psicoticism <- data_eng %>%
  select("X", "PSICOTICISMO1", "PSICOTICISMO2", "PSICOTICISMO3", "depreposparto") %>%
  gather(trimester, psicoticism, PSICOTICISMO1:PSICOTICISMO3, -depreposparto)


ggplot(tidy_psicoticism, aes(x = trimester, y = psicoticism, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Psicoticism", colour = "Postpartum Depression")+
  ggtitle('Psicoticism') +
  theme_bw() 
```

```{r}
tidy_IGS <- data_eng %>%
  select("X", "IGS1", "IGS2", "IGS3", "depreposparto") %>%
  gather(trimester, IGS, IGS1:IGS3, -depreposparto)


ggplot(tidy_IGS, aes(x = trimester, y = IGS, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Global Severity Index", colour = "Postpartum Depression")+
  ggtitle('Global Severity Index') +
  theme_bw() 
```

```{r}
tidy_SP <- data_eng %>%
  select("X", "SP1", "SP2", "SP3", "depreposparto") %>%
  gather(trimester, SP, SP1:SP3, -depreposparto)


ggplot(tidy_SP, aes(x = trimester, y = SP, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Positive symptoms total", colour = "Postpartum Depression")+
  ggtitle('Positive symptoms total') +
  theme_bw() 
```

```{r}
tidy_PSDI <- data_eng %>%
  select("X", "PSDI1", "PSDI2", "PSDI3", "depreposparto") %>%
  gather(trimester, PSDI, PSDI1:PSDI3, -depreposparto)


ggplot(tidy_PSDI, aes(x = trimester, y = PSDI, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Positive symptoms distress index", colour = "Postpartum Depression")+
  ggtitle('Positive symptoms distress index') +
  theme_bw() 
```

##Figure 3
###Hair cortisol levels differences (pg/mg) in each trimester between women with and without postpartum depression symptoms.

```{r}
tidy_LNcor <- data_eng %>%
  select("X", "LNCORTISOL1", "LNCORTISOL2", "LNCORTISOL3", "depreposparto") %>%
  gather(trimester, LNCORTISOL, LNCORTISOL1:LNCORTISOL3, -depreposparto)

tidy_LNcor$depreposparto <- as.factor(tidy_LNcor$depreposparto)

ggplot(tidy_LNcor, aes(x = trimester, y = LNCORTISOL, colour = depreposparto)) +
  stat_summary(fun.y = mean, geom = "point") + 
  stat_summary(fun.y = mean, geom = "line", size = 1, aes(group= factor(depreposparto))) + 
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.1, size = 0.5) + 
  labs(x = "Trimesters", y = "Hair cortisol levels differences", colour = "Postpartum Depression")+
  theme_bw() 
```

##Table 3
###Maternal hair cortisol levels and sex of the fetus.
```{r}
data_eng %>%
  group_by(SexFetalDico) %>%
  summarise(mean_cor1 = mean(LNCORTISOL1), sd_cor1 = sd(LNCORTISOL1), mean_cor2 = mean(LNCORTISOL2), sd_cor2 = sd(LNCORTISOL2), mean_cor3 = mean(LNCORTISOL3), sd_cor3 = sd(LNCORTISOL3))

t.test(data_eng$LNCORTISOL1 ~ data_eng$SexFetalDico, var.equal = T) 
t.test(data_eng$LNCORTISOL2 ~ data_eng$SexFetalDico, var.equal = T) 
t.test(data_eng$LNCORTISOL3 ~ data_eng$SexFetalDico, var.equal = T) 
```

##Multiple linear regression
```{r}
lm_ppd <- lm(epds ~ cortisol_tri1 + cortisol_tri2 + cortisol_tri3, data = pp_sad)
lm_ppd

summary(lm_ppd)
```

# EXTENSION

## Summary   
This extension has two parts. First, we look at how a MLR preforms when including all paper reported statistically significant variables. We do this to highlight that their findings are not externally valid or p-adjusted and therefore should not perform well in a model. As we expeceted, the colinearity between variables results in only one variable being signifincat in the MLR. Interestingly, the model still performs fairly well when applied to a test set, although, we really don't have a large enough dataset to validate this. The second approach we took was to extract a new feature combining all the cortisol data. We did this by calculating the area under the curve (AUC) and then fitting a probit regression to predict postpartum depression. This model performed admirably and could be a useful tool to predict postpartum depression, although, the specificity was rather low. 

## Multiple Linear Regression w/ all 'significant' variables 
### remember, there is high colinearity between these variables and p.values were unadjusted 

```{r}
lm_ppd2 <- lm(epds ~ cortisol_tri1 + cortisol_tri3 + depression_tri2 + depression_tri1 + depression_tri3 + employed + first_pregnancy + fetus_sex, data = pp_sad)
lm_ppd2

summary(lm_ppd2)


```

```{r}
#Performance
##### Train Test Split
library(caTools)
set.seed(101)
sample <- sample.split(pp_sad, SplitRatio = 0.75)
train <- subset(pp_sad, sample==T)
test <- subset(pp_sad, sample==F)

lm_fit <- lm(epds ~ cortisol_tri1 + cortisol_tri3 + depression_tri2 + depression_tri1 + depression_tri3 + employed + first_pregnancy + fetus_sex, data = train)

summary(lm_fit)

## RMSE
# Predict in-sample
predicted <- predict(lm_fit, test)
# Calculate RMSE
actual <- test$epds
sqrt(mean((predicted - actual)^2))


##predict in test model
test$lm_predict <- predict(lm_fit, test)

test$obs_PPD <- ifelse(test$epds < 10, "OB_No_PPD", "OB_PPD")
test$predict_PPD <- ifelse(test$lm_predict < 10, "No PPD", "PPD")

# Table
table(test$predict_PPD, test$obs_PPD)

library(pROC)

# ROC curve
roc_PPD <- roc(obs_PPD ~ lm_predict, data = test, percent = T)
print(roc_PPD)

#Drow ROC curve
plot.roc(roc_PPD, print.auc=TRUE, col="blue")

```

## Cortisol AUC feature and probit regression 

### Question: Can we create a feature from AUC of the cortisol trimesters to better predict of dichotomos postpartum depression diagnosis? 

### Method 
1. Fit curve or use **linear interpolation** of cortisol measurements over the three trimesters.   
2. Compute the integral of the fitted curve over the relevant time range (3 trimesters).  
3. Create a probit regression glm from cortisol AUC features to predict postpartum depression. 

### Example for AUC feature extraction **if** (we didn't) we fit a linear model. 
$$ AUC = \int_{0}^{6} (\beta x + C) \space dx \\ ==> 18B + 6C $$

### Load Tidy Dataset 
```{r}
load('../../data/tidy_data.Rdata')
pp_sad <- df_tidy %>% mutate(postpartum_depression=postpartum_depression-1)
``` 

### Calculating the AUC by **linear interpolation** and then using a probit fit to predict dichotomous postpartum depression.
```{r}
AUC <- vector(mode = 'numeric', length=44)
for (i in seq(1,44,1)){ 
  patient <- pp_sad %>% filter(row_number() == i)
  
  y <- c(patient$cortisol_tri1, patient$cortisol_tri2, patient$cortisol_tri3)
  x <- c(2, 5, 8) #months, what are the actual time frames we should use?
  
  AUC[i] <- AUC(x, y, method = "trapezoid", na.rm = FALSE) # this uses interpolation between our data points instead of lm fit
}

pp_sad <- pp_sad %>% mutate(Cortisol_AUC = AUC)

pp_sad %>% ggplot(aes(x=Cortisol_AUC, y=postpartum_depression)) + geom_point() + geom_smooth(method='glm', method.args=list(family=binomial(link="probit")))

pp_sad %>% ggplot(aes(x=Cortisol_AUC, group=postpartum_depression, fill=as.factor(postpartum_depression), alpha=0.25)) + geom_density()

```

```{r} 
pp_probit <- glm('postpartum_depression ~ Cortisol_AUC', family = binomial(link = 'probit'), data = pp_sad)

t.tst <- t.test(Cortisol_AUC ~ as.factor(postpartum_depression), data=pp_sad, na.rm=TRUE) 
t.tst

summary(pp_probit)

pp_sad <- pp_sad %>% mutate(pred = ifelse(plogis(predict(pp_probit, pp_sad)) > 0.5, 1, 0)) # predicted scores
TP <- pp_sad %>% filter(pred == 1 & postpartum_depression == 1.0) %>% nrow()
FP <- pp_sad %>% filter(pred == 1 & postpartum_depression == 0.0) %>% nrow()
TN <- pp_sad %>% filter(pred == 0 & postpartum_depression == 0.0) %>% nrow()
FN <- pp_sad %>% filter(pred == 0 & postpartum_depression == 1.0) %>% nrow()
sensitivity <- TP / (TP+FN)
specificity <- TN / (TN + FP)
truth.tab <- data.frame(TP,TN,FP,FN, specificity, sensitivity)
truth.tab
```
## Conclusion 
This extension shows value as a feature extraction method and resulted in a statistically significant predictive variable. 

# References 

[1] Caparros-Gonzalez, Rafael A., et al. "Hair Cortisol Levels, Psychological Stress and Psychopathological Symptoms as Predictors of Postpartum Depression." PLOS ONE, Public Library of Science, journals.plos.org/plosone/article/metrics?id=10.1371/journal.pone.0182817#citedHeader. 
