---
title: 'Final Project, P3: EDA'
author: "Nathaniel Evans, ALfonso Poire, Wei-Chun Lin"
date: "November 23, 2018"
output:
  html_document: 
    theme: flatly 
    highlight: kate
    toc: true
    toc_depth: 2
    toc_float : true
    smooth_scroll: true
    #number_sections : true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(GGally)

```



## Load Tidy Dataset 

```{r}
getwd()

load('../../data/tidy_data.Rdata')

pp_sad <- df_tidy # rename meaningful 

``` 


# Examine Raw Data 

```{r}
glimpse(pp_sad)

summary(pp_sad)
```


# Visualize 
```{r}
#ggpairs(pp_sad) # takes forever obviously
```


# Let your questions guiiiddddeeeee you


How does depression scores map to postpartum_depression labels? Might there be a better threshold cuttoff?
#the accuracy of the tables seem to be dependent on the trimester when the test is taken they are more accurate in the third trimester then the first this is to be expected. however the emans do not seem to vary much even in the third trimester so perhaps this is not a very accurate method regardless. trimester 2 seems to be the most infulential predictor. 
```{r}
#variation of depression scores in trimesters seperated by outcome of post partum depression or not 
ggplot(pp_sad,aes(x=depression_tri1))+
  geom_histogram(binwidth = 10,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()
ggplot(pp_sad,aes(x=depression_tri2))+
  geom_histogram(binwidth = 10,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()
ggplot(pp_sad,aes(x=depression_tri3))+
  geom_histogram(binwidth = 10,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()


```

```{r}
#same info shown in boxplots
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri1,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri2,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri3,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
```
#did the same data crunches on the cortosol levels in the different trimesters
#cortosol seems to be a better indecator than the tests but again not super different however the first trimester seems to show the biggest difference in the depressed verses not depressed groups 
```{r}
ggplot(pp_sad,aes(x=cortisol_tri1))+
  geom_histogram(binwidth = 50,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()
ggplot(pp_sad,aes(x=cortisol_tri2))+
  geom_histogram(binwidth = 50,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()
ggplot(pp_sad,aes(x=cortisol_tri3))+
  geom_histogram(binwidth = 50,col="white")+
  facet_wrap(~postpartum_depression)+
  theme_gray()
```

```{r}
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri1,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri2,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri3,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
```
#Aslso checked to see if there was relation between enployment status and cortosol levels or  depression scale score
```{r}
ggplot(pp_sad,aes(x=employed,y=cortisol_tri1,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=cortisol_tri2,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=cortisol_tri3,group=employed))+
  geom_boxplot()+
  theme_gray()
```
```{r}
ggplot(pp_sad,aes(x=employed,y=depression_tri1,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=depression_tri2,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=depression_tri3,group=employed))+
  geom_boxplot()+
  theme_gray()
```
#Aslso checked to see if there was relation between enployment type and cortosol levels or  depression scale score
```{r}
ggplot(pp_sad,aes(x=occupation,y=depression_tri1,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=depression_tri2,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=depression_tri3,group=occupation))+
  geom_boxplot()+
  theme_gray()
```
```{r}
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri1,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri2,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri3,group=occupation))+
  geom_boxplot()+
  theme_gray()
```
#sertain professions do seem to correlate to higher depression and cortosol levels but the sample size is very samall
#lastly checked occupation vs post partum depression and employment vs post partum depression
#this would seem to suggest some correlation between employment status and depression 
```{r}
ggplot(pp_sad,aes(x=occupation,y=postpartum_depression))+
  geom_col()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=postpartum_depression))+
  geom_col()+
  theme_gray()
```
#did the same comarison with wanted vs unwanted  and first pregnancy
```{r}
ggplot(pp_sad,aes(x=wanted_pregnancy,y=postpartum_depression))+
  geom_col()+
  theme_gray()
ggplot(pp_sad,aes(x=first_pregnancy,y=postpartum_depression))+
  geom_col()+
  theme_gray()
```
#sex of fetus vs depression 
```{r}
ggplot(pp_sad,aes(x=fetus_sex,y=postpartum_depression))+
  geom_col()+
  theme_gray()
```

#cortosol and depression relation across trimesters for to lm intresting how it changes from trimester to trimester becoming negative in the end
```{r}
ggplot(pp_sad,aes(x=depression_tri1,y=cortisol_tri1,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
ggplot(pp_sad,aes(x=depression_tri2,y=cortisol_tri2,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
ggplot(pp_sad,aes(x=depression_tri3,y=cortisol_tri3,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
```

Colinearlity between variables? what do we want to avoid including in our model?
```{r}
pp_sad%>%cor()
```

any obviously mislabeled or confusing trends etc etc
#having a bit of a hard time buying the over all result at this point. seems like there are enought other factors that also show some correlation with depression that saying cortosol is a good predictor is hard 


## Future Extension ... (I got bored)
This isn't really EDA, more of extension, but... I wonder if AUC of the cortisol points would be a better predictor of pp depression than just the point values. Would require fitting a curve, calculating the integral.... ORRRR could we use a curve to predict cortisol levels during pp? then use THAT metric to predict depression. 

$$ auc = \int_{0}^{6} (\beta x + C) \space dx \\ ==> 18B + 6C $$
```{r}
AUC <- vector(mode = 'numeric', length=44)
for (i in seq(1,44,1)){ 
  patient <- pp_sad %>% filter(row_number() == i)
  
  y <- c(patient$cortisol_tri1, patient$cortisol_tri2, patient$cortisol_tri3)
  x <- c(0, 3, 6) #months, what are the actual time frames we should use?
  
  df <- data.frame( 'x' = x, 'y' = y )
  
  lm <- glm(data=df, formula= 'y ~ x') # what do we expect our data to look like? what kind of fitted model should we expect? How do we prevent under/over fitting ? 
  
  #plt <- ggplot(data=df, aes(x=x, y=y)) + geom_point() + geom_smooth(method='glm')
  #print(plt)
  
  B <- lm$coefficients['x']
  C <- lm$coefficients['(Intercept)']
  
  this_AUC <- 18*B + 6*C 
  
  AUC[i] <- this_AUC
  
  #print(AUC)
}



pp_sad <- pp_sad %>% mutate(Cortisol_AUC = AUC, postpartum_depression=postpartum_depression)

pp_sad %>% ggplot(aes(x=Cortisol_AUC, y=postpartum_depression)) + geom_point() + geom_smooth(method='glm', method.args=list(family=binomial(link="probit")))

pp_sad %>% ggplot(aes(x=Cortisol_AUC, group=postpartum_depression, color=postpartum_depression)) + geom_density()

```

