---
title: 'Final Project, P3: EDA'
author: "Nathaniel Evans, ALfonso Poire, Wei-Chun Lin"
date: "November 23, 2018"
output:
  html_document: 
    theme: flatly 
    highlight: kate
    toc: true
    toc_depth: 2
    toc_float : true
    smooth_scroll: true
    #number_sections : true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(GGally)

```



## Load Tidy Dataset 

```{r}
getwd()

load('../../data/tidy_data.Rdata')

pp_sad <- df_tidy # rename meaningful 

``` 


# Examine Raw Data 

```{r}
glimpse(pp_sad)

summary(pp_sad)
```


# Visualize 
```{r}
#ggpairs(pp_sad) # takes forever obviously
```


# Let your questions guiiiddddeeeee you


How does depression scores map to postpartum_depression labels? Might there be a better threshold cuttoff?
#the accuracy of the tables seem to be dependent on the trimester when the test is taken they are more accurate in the third trimester then the first this is to be expected. however the emans do not seem to vary much even in the third trimester so perhaps this is not a very accurate method regardless. trimester 2 seems to be the most infulential predictor. 
```{r}
#variation of depression scores in trimesters seperated by outcome of post partum depression or not 

pp_sad %>% ggplot() + geom_density(aes(x=depression_tri1, fill='tri1', alpha = 0.3)) + geom_density(aes(x=depression_tri2, fill='tri2', alpha = 0.3)) + geom_density(aes(x=depression_tri3, fill='tri3', alpha = 0.3)) + facet_wrap(~postpartum_depression) + ggtitle('trimester depression populations by postpartum outcome') + xlab('depression score')

pp_sad %>% ggplot() + geom_density(aes(x=depression_tri1, group=as.factor(postpartum_depression), fill=postpartum_depression, alpha=0.25))


pp_sad %>% ggplot() + geom_density(aes(x=depression_tri2, group=as.factor(postpartum_depression), fill=as.factor(postpartum_depression), alpha = 0.25))


pp_sad %>% ggplot() + geom_density(aes(x=depression_tri3, group=postpartum_depression, fill=as.factor(postpartum_depression), alpha = 0.25))

# I like this! But I condensed everything into one graph to make it easier to compare
#ggplot(pp_sad,aes(x=depression_tri1))+
#  geom_histogram(binwidth = 10,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()
#ggplot(pp_sad,aes(x=depression_tri2))+
#  geom_histogram(binwidth = 10,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()
#ggplot(pp_sad,aes(x=depression_tri3))+
#  geom_histogram(binwidth = 10,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()


```

```{r}
pp_sad %>% ggplot() + geom_violin(aes(x=postpartum_depression,y=depression_tri1,group=postpartum_depression, fill='tri1', alpha=0.3)) + geom_violin(aes(x=postpartum_depression,y=depression_tri2,group=postpartum_depression, fill='tri2', alpha=0.3)) + geom_violin(aes(x=postpartum_depression,y=depression_tri3,group=postpartum_depression, fill='tri3', alpha=0.3))

#same info shown in boxplots
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri1,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri2,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=depression_tri3,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
```
#did the same data crunches on the cortosol levels in the different trimesters
#cortosol seems to be a better indecator than the tests but again not super different however the first trimester seems to show the biggest difference in the depressed verses not depressed groups 
```{r}
pp_sad %>% ggplot() + geom_density(aes(x=cortisol_tri1, fill='tri1', alpha = 0.3)) + geom_density(aes(x=cortisol_tri2, fill='tri2', alpha = 0.3)) + geom_density(aes(x=cortisol_tri3, fill='tri3', alpha = 0.3)) + facet_wrap(~postpartum_depression) + ggtitle('trimester cortisol populations by postpartum outcome') + xlab('cortisol level')

pp_sad %>% ggplot() + geom_density(aes(x=cortisol_tri1, group=as.factor(postpartum_depression), fill=postpartum_depression, alpha=0.25))


pp_sad %>% ggplot() + geom_density(aes(x=cortisol_tri2, group=as.factor(postpartum_depression), fill=as.factor(postpartum_depression), alpha = 0.25))


pp_sad %>% ggplot() + geom_density(aes(x=cortisol_tri3, group=postpartum_depression, fill=as.factor(postpartum_depression), alpha = 0.25))

# same as above 
#ggplot(pp_sad,aes(x=cortisol_tri1))+
#  geom_histogram(binwidth = 50,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()
#ggplot(pp_sad,aes(x=cortisol_tri2))+
#  geom_histogram(binwidth = 50,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()
#ggplot(pp_sad,aes(x=cortisol_tri3))+
#  geom_histogram(binwidth = 50,col="white")+
#  facet_wrap(~postpartum_depression)+
#  theme_gray()
```

```{r}
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri1,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri2,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=postpartum_depression,y=cortisol_tri3,group=postpartum_depression))+
  geom_boxplot()+
  theme_gray()




```
#Aslso checked to see if there was relation between enployment status and cortosol levels or  depression scale score
```{r}
ggplot(pp_sad,aes(x=employed,y=cortisol_tri1,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=cortisol_tri2,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=cortisol_tri3,group=employed))+
  geom_boxplot()+
  theme_gray()
```
```{r}
ggplot(pp_sad,aes(x=employed,y=depression_tri1,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=depression_tri2,group=employed))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=depression_tri3,group=employed))+
  geom_boxplot()+
  theme_gray()
```
#Aslso checked to see if there was relation between enployment type and cortosol levels or  depression scale score
```{r}
ggplot(pp_sad,aes(x=occupation,y=depression_tri1,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=depression_tri2,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=depression_tri3,group=occupation))+
  geom_boxplot()+
  theme_gray()
```
```{r}
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri1,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri2,group=occupation))+
  geom_boxplot()+
  theme_gray()
ggplot(pp_sad,aes(x=occupation,y=cortisol_tri3,group=occupation))+
  geom_boxplot()+
  theme_gray()
```
#Certain professions do seem to correlate to higher depression and cortisol levels but the sample size is very small
#Lastly, checked occupation vs post partum depression and employment vs post partum depression
#this would seem to suggest some correlation between employment status and depression 
```{r}
ggplot(pp_sad,aes(x=occupation,y=postpartum_depression))+
  geom_col()+
  theme_gray()
ggplot(pp_sad,aes(x=employed,y=postpartum_depression))+
  geom_col()+
  theme_gray()
```
#did the same comparison with wanted vs unwanted  and first pregnancy
```{r}
ggplot(pp_sad,aes(x=wanted_pregnancy,y=postpartum_depression))+
  geom_col()+
  theme_gray()
ggplot(pp_sad,aes(x=first_pregnancy,y=postpartum_depression))+
  geom_col()+
  theme_gray()
```
#sex of fetus vs depression 
```{r}
ggplot(pp_sad,aes(x=fetus_sex,y=postpartum_depression))+
  geom_col()+
  theme_gray()


```

#cortosol and depression relation across trimesters for to lm intresting how it changes from trimester to trimester becoming negative in the end
```{r}
ggplot(pp_sad,aes(x=depression_tri1,y=cortisol_tri1,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
ggplot(pp_sad,aes(x=depression_tri2,y=cortisol_tri2,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
ggplot(pp_sad,aes(x=depression_tri3,y=cortisol_tri3,color=postpartum_depression))+geom_point()+geom_smooth(method = "lm", se = FALSE)
```

Colinearlity between variables? what do we want to avoid including in our model?
```{r}
# we should really only check continuous variables 

.cor <- pp_sad %>% select(starts_with('cortisol'), starts_with('depression'), age, epds, education_level, fetus_sex, previous_miscarriage) %>% cor()
library(reshape2)
melted_sad <- melt(.cor)

ggplot(data = melted_sad, aes(x=Var1, y=Var2, fill=value)) + geom_tile() +  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +  theme(axis.text.x = element_text(angle = 60, vjust = 1, 
    size = 12, hjust = 1))
```

any obviously mislabeled or confusing trends etc etc
#having a bit of a hard time buying the over all result at this point. seems like there are enough other factors that also show some correlation with depression that saying cortisol is a good predictor is hard. From what I recall, PP depression does have psycological or social predictors but there haven't been any biological predictors, which was what this study was trying to extend. 


