---
title: "Data_Delivery"
author: "Nathaniel Evans, Win-Chun Lin, Alfonso Poire"
date: "November 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(ggplot2)
library(reticulate)
library(haven)
library(codebook)
library(knitr)
library(rmarkdown)
```

# MATH 630 GROUP PROJECT
## PART 2 : DATA DELIVERY 
### GROUP 7 : Nate Evans, Win-Chun Lin, Alfonso Poire 
### PAPER: [Hair cortisol levels, psychological stress and psychopathological symptoms as predictors of postpartum depression](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0182817) 
### DATA: [sav file](https://figshare.com/articles/PPD_hairCortisol_PlosOne_sav/5255848/1) 
### TIDY DATA: [csv file](./../../data/tidy_daya.csv)
### CODE BOOK: [knitted html](./group7_codebook.html)

#### The purpose of this R markdown script is to produce the following components, as detailed in the Ellis & Leek paper, with respect to our paper. For all the links to work, it is best to clone the the [github repository](https://github.com/OHSU-Math630/FINAL_PROJECT) and access this script within the cloned file strucutre (located: FINAL_PROJECT/postpartum_depression_predicition/P2/group7_DataDelivery.Rmd). 

1. The raw data : <span style="color: red"> This is provided in the link above, and/or as the .sav file uploaded to sakai </span>
2. A tidy data set (Wickham 2014) : <span style="color: red"> This will be proided in the link above as well as uploaded to sakai, depending on file structure, the link may be broken </span>
3. A code book describing each variable and its values in the tidy data set. [do this for all variables that will be part of your replication analyses] <span style="color: red"> This is in the form of a knitted html file, which can be found in the native file structure linked above or uploaded to sakai </span>
4. An explicit and exact recipe you used to go from 1 -> 2,3.  <span style="color: red"> This rmd script serves as the recipe from 1 -> 2,3 </span>


#Generating a tidy dataset 
#### The first step we took to produce a tidy dataset was to translate the variable names from english to spanish, to do this, we first attempted a programatic method; Using the google translate API, each variable name was attempted to translate from Spanish to English. 

```{r import data, convert to csv}
path = "./../../data/PPD_hairCortisol_PlosOne_.sav"

dataset= read_sav(path) #meta data dropped? 

write.csv(dataset, file = "./../../data/data_span.csv")
```

```{python}
import googletrans as gt 
import pandas as pd

data = pd.read_csv('./../../data/data_span.csv',encoding = "ISO-8859-1")

trans = gt.Translator()

with open("./../../data/data_programatic_translation.csv", 'w') as f: 
  f.write('original, translated\n')
  for old_col in data.columns.values:  
    new_col = trans.translate(old_col, src = 'spanish', dest='en').text
    f.write(old_col + ',' + new_col + '\n')
    
    data.rename(columns = {old_col:new_col}, inplace = True)
  
data.to_csv( "./../../data/data_eng.csv" )
```

#### This produces two files, first, a new dataset csv with the header names translated (attempted) and two, a csv documenting the original variable name and the translated variable name. </br> 

[english csv](./../../data/data_eng.csv)  
[translation mapping](./../../data/data_programatic_translation.csv)

### ALL VARIABLES WE NEED TO INCLUDE IN OUR DATASET
#### These listed variables are those required to reproduce the paper plots or extend the study. 

![Paper Variables](https://journals.plos.org/plosone/article/figure/image?size=large&id=info:doi/10.1371/journal.pone.0182817.t001) 

###We will try to manually review, translate and generate a [codebook](https://cran.r-project.org/web/packages/codebook/codebook.pdf) (Good tutorial: [github](https://github.com/rubenarslan/codebook)) in the following steps. 

## ALL of our raw data are numerics, where is the codebook mapping categorical -> numeric ??? like 'profession' 

```{r} 
df = read.csv("./../../data/data_eng.csv", header = TRUE)
#df %>% variable.names()

# THIS IS WHERE WE CAN DROP UNUNSED VARIABLES, MUTATE NAMES AND ADD META DATA DESCRIBING VARIABLES (variable attribute data)

# There are 29 rows of all NA variables that were brought in from SPSS empty rows 
before <- df %>% nrow() 
df <- df %>% na.omit() 
after <- df %>% nrow() 
print(paste('rows dropped:', after-before))

df <- df %>% mutate('age' = Age)
attr(df$age, "Description") <- "The age of the mother"

df <- df %>% mutate("nationality" = NationalityDicotomica)
attr(df$NationalityDicotomica, 'Description') <- "The patient's nationality or country of origin at the time of pregnancy"

#df <- df %>% mutate('Marital_Status' = ) # CANT FIND THIS ONE

df <- df %>% mutate('employed' = EmploymentSituationDico)
attr(df$employed, 'Description') <- "Employment situation; is or is not employed at time of pregnancy"

df <- df %>% mutate('occupation' = Profession)
attr(df$occupation, "Description") <- 'If employed, what was the mothers occupation at the time of pregnancy' 

df <- df %>% mutate("education_level" = Level_Studies) 
attr(df$education_level, 'Description') <- "The highest level of education that the mother had received at the time of pregnancy/study"

df <- df %>% mutate('sport' = Sport)
attr(df$sport, 'Description') <- 'tracking the physical activity of the mother; whether the mother played a sport (during pregnancy) [y/n]' # ???

df <- df %>% mutate('pet' = PetDicotomica)
attr(df$pet, 'Description') <- 'designation for mothers with pets at time of pregnancy' 

df <- df %>% mutate('dyed_hair' = Appearance)
attr(df$dyed_hair, "Description") <- 'designation for mothers with dyed hair during pregnancy'  # I THINK this is right, need to check

df <- df %>% mutate('first_pregnancy' = FirstPregnancy)
attr(df$first_pregnancy, 'Description') <- 'primiparous; designation for mothers who are having a child for the first time'

df <- df %>% mutate('wanted_pregnancy' = WantedPregnancyDico)
attr(df$wanted_pregnancy, 'Description') <- 'Designation for mothers who desired the pregnancy (planned?)'


df <- df %>% mutate('pregnancy_method' = PregnancyMethodDico)
attr(df$pregnancy_method, 'Description') <- 'designation for  *spontaneous* method of fertilization and *Fertility Treatment* or artificial insemination' # PC term? also, lol on "spontaneous" 

df <- df %>% mutate('previous_miscarriage' = PreviousMiscarriagesDico)
attr(df$previous_miscarriage, 'Description') <- 'Designation for mothers who have had a miscarriage prior to the current pregnancy' 

# df <- df %>% mutate('delivery' = ) # CANT FIND DELIVERY AND LABOR

#df <- mutate('delivery_antisthesia' = )    # Not sure what this one is 

df <- df %>% mutate('fetus_sex' = SexFetalDico)
attr(df$fetus_sex, "Description") <- 'the sex of the fetus' # numeric mapping ???  

# -------- DEPRESSION METRICS 

df <- df %>% mutate('postpartum_depression' = depreposparto)
attr(df$postpartum_depression, 'Description') <- "Diagnosis of post-partum depression in the mother" ### ???? more?


df <- df %>% mutate('depression_tri1' = DEPRESSION1)
df <- df %>% mutate('depression_tri2' = DEPRESSION2)
df <- df %>% mutate('depression_tri3' = DEPRESSION3)
attr(df$depression_tri1, 'Description') <- 'antenatal depression during the 1st trimester.'
attr(df$depression_tri2, 'Description') <- 'antenatal depression during the 2nd trimester.'
attr(df$depression_tri3, 'Description') <- 'antenatal depression during the 3rd trimester.'

df <- df %>% mutate('epds' = EPDS)
attr(df$epds, 'Description') <- 'continuous metric of postnatal depression'  #measurement name, type? 

# --------- CORTISOL metrics 

df <- df %>% mutate('cortisol_tri1' = Cortisol1) #WHAT IS LNCORTISOL1 ????? - Two metrics for coritsol??
df <- df %>% mutate('cortisol_tri2' = Cortisol2)
df <- df %>% mutate('cortisol_tri3' = Cortisol3)
attr(df$cortisol_tri1, 'Description') <- 'Cortisol levels in mothers during pregnancy during the 1st trimester' # units??? look into lncortisol variable, which one should we be using? 
attr(df$cortisol_tri2, 'Description') <- 'Cortisol levels in mothers during pregnancy during the 2nd trimester'
attr(df$cortisol_tri3, 'Description') <- 'Cortisol levels in mothers during pregnancy during the 3rd trimester'

# ---------- SELECT 

df_tidy <- df %>% select(age, education_level, nationality, postpartum_depression, employed, occupation, sport, sport, pet, dyed_hair, first_pregnancy, wanted_pregnancy, pregnancy_method, previous_miscarriage, fetus_sex, epds, depression_tri1, depression_tri2, depression_tri3, cortisol_tri1, cortisol_tri2, cortisol_tri3 )

#glimpse(df_tidy)

write.csv(df_tidy, file='./../../data/tidy_data.csv') # this might drop Description attribute 
save(df_tidy, file='./../../data/tidy_data.Rdata') # this way we can always load it back as is. 

```

## The tidy dataset can be found in the data directory, linked [here](./../../data/tidy_data.csv)

```{r}
knitr::opts_chunk$set(warning = TRUE, message = TRUE, echo = FALSE) #this has to go before the codebook() call, make sure echo=FALSE is included. 
my_codebook <- codebook(df_tidy) #my_codebook # (TAKES FOREVER TO KNIT) This will produce a codebook, but we need to have this knitted in it's own html file and not echo the code chunks... 
knitr::opts_chunk$set(warning = TRUE, message = TRUE, echo = TRUE)

fh<-file("tmp.Rmd")
writeLines(my_codebook, fh)
close(fh)

render(input='tmp.Rmd', output_file = 'group7_codebook.html', output_dir = getwd() )
```

### The codebook is a knitted html file in the P2 folder and is linked [here](./group7_codebook.html)
#### The dataframe __df_tidy__ has an attribute "Description" that can be accessed with fn: attribute(df$variable, "Description")



















#REFERENCES 

[1] codebook package: 

Preprint
Arslan, R. C. (2018). How to automatically generate rich codebooks from study metadata. doi:10.31234/osf.io/5qc6h

Zenodo
Arslan, R. C. (2018). Automatic codebooks from survey metadata (2018). URL https://github.com/rubenarslan/codebook. DOI





