---
title: "Data_Delivery"
author: "Nathaniel Evans, Win-Chun Lin, Alfonso Poire"
date: "November 11, 2018"
output: 
  html_document: 
    theme: flatly 
    highlight: kate
    toc: true
    toc_depth: 2
    toc_float : true
    smooth_scroll: true
    #number_sections : true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(ggplot2)
library(reticulate)
library(haven)
library(codebook)
library(knitr)
library(rmarkdown)
library(dataMaid) #alternative codebook 
```

#Overview
## PART 2 : DATA DELIVERY 

### PAPER: [Hair cortisol levels, psychological stress and psychopathological symptoms as predictors of postpartum depression](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0182817) 
### DATA: [sav file](https://figshare.com/articles/PPD_hairCortisol_PlosOne_sav/5255848/1) 
### TIDY DATA: [csv file](./../../data/tidy_daya.csv)
### CODE BOOK: [knitted html](./group7_codebook.html)

#### The purpose of this R markdown script is to produce the following components, as detailed in the Ellis & Leek paper, with respect to our paper. For all the links to work, it is best to clone the the [github repository](https://github.com/OHSU-Math630/FINAL_PROJECT) and access this script within the cloned file strucutre (located: FINAL_PROJECT/postpartum_depression_predicition/P2/group7_DataDelivery.Rmd). 

1. The raw data : <span style="color: red"> This is provided in the link above, and/or as the .sav file uploaded to sakai </span>
2. A tidy data set (Wickham 2014) : <span style="color: red"> This will be proided in the link above as well as uploaded to sakai, depending on file structure, the link may be broken </span>
3. A code book describing each variable and its values in the tidy data set. [do this for all variables that will be part of your replication analyses] <span style="color: red"> This is in the form of a knitted html file, which can be found in the native file structure linked above or uploaded to sakai </span>
4. An explicit and exact recipe you used to go from 1 -> 2,3.  <span style="color: red"> This rmd script serves as the recipe from 1 -> 2,3 </span>

#Generating a tidy dataset 
## importing data, converting to csv 

```{r import data, convert to csv}
path = "./../../data/PPD_hairCortisol_PlosOne_.sav"

dataset= read_sav(path) #meta data dropped? Yes. Use SPSS to access

write.csv(dataset, file = "./../../data/data_span.csv")
```

## Programatic Translation
#### The first step we took to produce a tidy dataset was to translate the variable names from english to spanish, to do this, we first attempted a programatic method; Using the google translate API, each variable name was attempted to translate from Spanish to English. 

#### Result: Not terribly useful, but a valuable concept. We end up manually reviewing and translating it all anyways. 

```{python}
import googletrans as gt 
import pandas as pd

data = pd.read_csv('./../../data/data_span.csv',encoding = "ISO-8859-1")

trans = gt.Translator()

with open("./../../data/data_programatic_translation.csv", 'w') as f: 
  f.write('original, translated\n')
  for old_col in data.columns.values:  
    new_col = trans.translate(old_col, src = 'spanish', dest='en').text
    f.write(old_col + ',' + new_col + '\n')
    
    data.rename(columns = {old_col:new_col}, inplace = True)
  
data.to_csv( "./../../data/data_eng.csv" )
```

#### This produces two files, first, a new dataset csv with the header names translated (attempted) and two, a csv documenting the original variable name and the translated variable name. </br> 

[english csv](./../../data/data_eng.csv)  
[translation mapping](./../../data/data_programatic_translation.csv)


## VARIABLES TO INCLUDE
#### These listed variables are those required to reproduce the paper plots or extend the study. 

![Paper Variables](https://journals.plos.org/plosone/article/figure/image?size=large&id=info:doi/10.1371/journal.pone.0182817.t001) 

##Data Cleaning
###We will try to manually review, translate and generate a [codebook](https://cran.r-project.org/web/packages/codebook/codebook.pdf) (Good [tutorial](https://github.com/rubenarslan/codebook)) in the following steps. 

#### ALL of our raw data are numerics, there is codebook mapping categorical -> numeric in __some__ of the variables if the .sav file is interograted from SPSS. We have tried to include as much of this information as possible, but the meta data was dropped when the sav file is imported into R, so much was lost. 

#### Summary: dropped na rows (29), renamed variables and added attribute 'Description' which can be found in the **summary statistics** tab of our codebook. 

```{r} 
df = read.csv("./../../data/data_eng.csv", header = TRUE)
df %>% variable.names()

# THIS IS WHERE WE CAN DROP UNUNSED VARIABLES, MUTATE NAMES AND ADD META DATA DESCRIBING VARIABLES (variable attribute data)

# There are 29 rows of all NA variables that were brought in from SPSS empty rows 
before <- df %>% nrow() 
df <- df %>% na.omit() 
after <- df %>% nrow() 
print(paste('rows dropped:', after-before))

df <- df %>% mutate('age' = Age)
attr(df$age, "shortDescription") <- "The age of the mother"

df <- df %>% mutate("nationality" = NationalityDicotomica)
attr(df$nationality, 'shortDescription') <- "The patient's nationality or country of origin at the time of pregnancy
[1 = Spain, 2 = Argentina, 3 = Chile, 4 = Morocco, 5 = Romania, 6 = Germany, 7 = Russia, 8 = Ecuador, 9 = Peru, 10 = Paraguay]"

#df <- df %>% mutate('Marital_Status' = ) # CANT FIND THIS ONE

df <- df %>% mutate('employed' = EmploymentSituationDico)
attr(df$employed, 'shortDescription') <- "Employment situation; is or is not employed at time of pregnancy [1 = unemployed, 2 = full-time job, 3 = half-day job, 4 = student, 5 = work and study"

df <- df %>% mutate('occupation' = Profession)
attr(df$occupation, "shortDescription") <- 'If employed, what was the mothers occupation at the time of pregnancy [1 = unemployment, 2 = Head of nurse, 3 = Doctor, 4 = Nurse, 5 = housewife, 6 = teacher, 7 = management, 8 = dental clinic, 9 = banking, 10 = untrained, 11 = psychologist, 12 = engineer'

df <- df %>% mutate("education_level" = Level_Studies) 
attr(df$education_level, 'shortDescription') <- "The highest level of education that the mother had received at the time of pregnancy/study :
1 = primary
2 = secondary
3 = university students
4 = without studies
"
df <- df %>% mutate('sport' = Sport)
attr(df$sport, 'shortDescription') <- 'tracking the physical activity of the mother; whether the mother played a sport (during pregnancy) [y/n] :
0 = No
1 = Yes' 

df <- df %>% mutate('pet' = PetDicotomica)
attr(df$pet, 'shortDescription') <- 'designation for mothers with pets at time of pregnancy' 

df <- df %>% mutate('dyed_hair' = Appearance)
attr(df$dyed_hair, "shortDescription") <- 'designation for mothers with dyed hair during pregnancy :
0 = dyed
1 = natural'
# I THINK this is right, need to check

df <- df %>% mutate('first_pregnancy' = FirstPregnancy)
attr(df$first_pregnancy, 'shortDescription') <- 'primiparous; designation for mothers who are having a child for the first time'

df <- df %>% mutate('wanted_pregnancy' = WantedPregnancyDico)
attr(df$wanted_pregnancy, 'shortDescription') <- 'Designation for mothers who desired the pregnancy (planned?)'


df <- df %>% mutate('pregnancy_method' = PregnancyMethodDico)
attr(df$pregnancy_method, 'shortDescription') <- 'designation for  *spontaneous* method of fertilization and *Fertility Treatment* or artificial insemination : 
1 = spontaneously and normally
2 = in vitro fertilization
3 = artificial insemination'

# PC term? also, lol on "spontaneous" 

df <- df %>% mutate('previous_miscarriage' = PreviousMiscarriagesDico)
attr(df$previous_miscarriage, 'shortDescription') <- 'Designation for mothers who have had a miscarriage prior to the current pregnancy' 

# df <- df %>% mutate('delivery' = ) # CANT FIND DELIVERY AND LABOR

#df <- mutate('delivery_antisthesia' = )    # Not sure what this one is 

df <- df %>% mutate('fetus_sex' = SexFetalDico)
attr(df$fetus_sex, "shortDescription") <- 'the sex of the fetus' # numeric mapping ???  

# -------- DEPRESSION METRICS 

df <- df %>% mutate('postpartum_depression' = depreposparto)
attr(df$postpartum_depression, 'shortDescription') <- "Diagnosis of post-partum depression in the mother: 
1 = sindepre?
2 = condepre?" ### ???? more ?


df <- df %>% mutate('depression_tri1' = DEPRESSION1)
df <- df %>% mutate('depression_tri2' = DEPRESSION2)
df <- df %>% mutate('depression_tri3' = DEPRESSION3)
attr(df$depression_tri1, 'shortDescription') <- 'antenatal depression during the 1st trimester.'
attr(df$depression_tri2, 'shortDescription') <- 'antenatal depression during the 2nd trimester.'
attr(df$depression_tri3, 'shortDescription') <- 'antenatal depression during the 3rd trimester.'

df <- df %>% mutate('epds' = EPDS)
attr(df$epds, 'shortDescription') <- 'continuous metric of postnatal depression'  #measurement name, type? 

# --------- CORTISOL metrics 
#the data set also contains a ln cortosol metic that is a log naturalised version of the normal cortosol metric
df <- df %>% mutate('cortisol_tri1' = Cortisol1) #WHAT IS LNCORTISOL1 ????? - Two metrics for coritsol??
df <- df %>% mutate('cortisol_tri2' = Cortisol2)
df <- df %>% mutate('cortisol_tri3' = Cortisol3)
attr(df$cortisol_tri1, 'shortDescription') <- 'Cortisol levels in mothers during pregnancy during the 1st trimester' # units??? look into lncortisol variable, which one should we be using? 
attr(df$cortisol_tri2, 'shortDescription') <- 'Cortisol levels in mothers during pregnancy during the 2nd trimester'
attr(df$cortisol_tri3, 'shortDescription') <- 'Cortisol levels in mothers during pregnancy during the 3rd trimester'

# ---------- SELECT 

df_tidy <- df %>% select(age, education_level, nationality, postpartum_depression, employed, occupation, sport, sport, pet, dyed_hair, first_pregnancy, wanted_pregnancy, pregnancy_method, previous_miscarriage, fetus_sex, epds, depression_tri1, depression_tri2, depression_tri3, cortisol_tri1, cortisol_tri2, cortisol_tri3)

#glimpse(df_tidy)

write.csv(df_tidy, file='./../../data/tidy_data.csv') # this might drop Description attribute 
save(df_tidy, file='./../../data/tidy_data.Rdata') # this way we can always load it back as is. 

```

## Tidy Data
### The tidy dataset can be found in the data directory, linked [here](./../../data/tidy_data.csv)

## Producing Codebook
### Adding information attribute to our tidy data to map numerics to categoricals
```{r}

numeric_mapping <- '# Numeric Mapping
Primer embarazo / First Pregnancy:
0 = No
1 = Yes

Tipo Embarazo / Pregnancy Method:
1 = spontaneously and normally
2 = in vitro fertilization
3 = artificial insemination

Nivel Estudios / level of education
1 = primary
2 = secondary
3 = university students
4 = without studies

Laboral / employee situation
1 = unemployed
2 = full-time job
3 = half-day job
4 = student
5 = work and study

Aspecto Pelo/ Hair Appearance
0 = dyed
1 = natural

Deporte / Sport
0 = No
1 = Yes

Profesión / Profession
1 = unemployment
2 = Head of nurse
3 = Doctor
4 = Nurse
5 = housewife
6 = teacher
7 = management
8 = dental clinic
9 = banking
10 = untrained
11 = psychologist
12 = engineer

Origen / Nationality
1 = Spain
2 = Argentina
3 = Chile
4 = Morocco
5 = Romania
6 = Germany
7 = Russia
8 = Ecuador
9 = Peru
10 = Paraguay

depreposparto / depress post partum
1 = sindepre?
2 = condepre?'

attr(df_tidy, 'numeric mapping information') <- numeric_mapping

```

### Generate a knitted codebook
```{r, message=FALSE}
# datamaid codebook
dataMaid::makeCodebook(df_tidy, replace=TRUE, codebook=TRUE, output='html')


#codebook codebook

df_tmp <- detect_scales(df_tidy) # loook for scalings between variables

knitr::opts_chunk$set(warning = TRUE, message = TRUE, echo = FALSE) #this has to go before the codebook() call, make sure echo=FALSE is included. 
my_codebook <- codebook(df_tmp) #my_codebook # (TAKES FOREVER TO KNIT) This will produce a codebook, but we need to have this knitted in it's own html file and not echo the code chunks... 
knitr::opts_chunk$set(warning = TRUE, message = TRUE, echo = TRUE)

fh<-file("tmp.Rmd")
writeLines(paste(numeric_mapping,my_codebook), fh)
close(fh)

rmarkdown::render(input='tmp.Rmd', output_file = 'group7_codebook.html', output_dir = getwd() )
```
## Codebook(S) 
### We tried both the codebook and datamaid packages for autogenerating a codebook and both had benefits and shortfalls. In the end, however, I think the datamaid is a bit more user friendly and displayed the variables in a cleaner way. The datamaid codebook is the version that will be submitted to sakai. 

####The **dataMaid package** [codebook](./codebook_df_tidy.html)
####The **codebook package** [codebook](./group7_codebook.html)
</br> 
#### The dataframe __df_tidy__ has an attribute "shortDescription" that can be accessed with fn: attribute(df$variable, "Description")



















#REFERENCES 

[1] codebook package: 

Preprint
Arslan, R. C. (2018). How to automatically generate rich codebooks from study metadata. doi:10.31234/osf.io/5qc6h

Zenodo
Arslan, R. C. (2018). Automatic codebooks from survey metadata (2018). URL https://github.com/rubenarslan/codebook. DOI





